Evaluating `Graphical Perception' with CNNs

### 2018-03-29 PAPER/02_experimental_setup.tex
Add just lil.

We conduct quantitative experiments to measure how different convolutional neural networks ``perceive'' low-level visual encodings such as positions, angles, curvature, and lengths. We treat each network as a subject in a between-subjects experiment to evaluate supervised learning algorithms. For this, we formulate any estimation of single quantities or relations between multiple quantities as a logistic regression problem with single or multiple outputs between $0$ and $1$. The output indicates the percentage in regards to the degrees of freedom of the individual experiment but all experiments have different sets of parameters. As baseline, we use a standard multilayer perceptron (MLP) without the generation of additional feature maps through convolutions (Fig.~\ref{fig:classifiers}).
For our experiments, we use a series of single factor between-subject designs with the factor \emph{classifier} (MLP, LeNet, VGG19, Xception, VGG19+ImageNet, Xception+ImageNet). Each experiment includes a training and a testing phase on data without any overlap (Section~\ref{sec:data}). All parameters such as network hyperparameters, optimization methods, and stopping conditions are fixed (Section~\ref{sec:classifiers}). However, since the classifiers are of different complexity, the number of trainable parameters changes (Table~\ref{tab:parameters}). We define a series of hypotheses prior to each experiment.

@@

\subsection{Data}
\label{sec:data}
We create all visualizations as parametrized rasterized images without interpolation (except for the anti-alias experiment). The number of parameters differs per experiment as summarized in Table~\ref{tab:encoding_parameters} and section~\ref{sec:parametrizations}. 
\\~\\
\noindent\textbf{Noise.} We add subtle random noise ($0.05$) to each pixel to introduce additional variation.
\\~\\
\noindent\textbf{Training/Validation/Test Splits.} We specify the size of each split set as follows: 60,000 training images, 20,000 validation images, and 20,000 test images. We then randomly add parameterized visualizations to the sets while guaranteeing that each set is disjunct from each other in terms of encoded variables. This eliminates leakage during training and evaluation. We also scale each set independently: images to the range of $-0.5$ to $0.5$ and labels to the range of $0.0$ to $1.0$. 
\\~\\
\noindent\textbf{Cross Validation.} For reproducibility, we perform repeated random sub-sampling validation, also known as Monte Carlo cross-validation, during our experiments. We run every experiment seperately multiple times (at least 4, up to 12) and randomly select (without replacement) the $60\%$ of our data as training data, $20\%$ as validation, and $20\%$ as test. Our large data sample size of 100,000 guarantees that every single observation of our parameterizations will be selected at least once (excluding noise patterns). Finally, we average the results over the runs.
\\~\\
\noindent\textbf{Intra-classifier Variability.} We also evaluate classifiers previously trained with one visualization on the same type of visualizations with different parameters by decreasing and increasing the variability of the generated images.
\\~\\
\noindent\textbf{Multi Classifiers.} Some experiments compare different types of visual encodings. We first train and evaluate each classifier on one type of encoding but we also train the classifiers on a random selection of multiple different types. These scenarios afflict the optimization process to find global minima and result in networks with more flexible knowledge with the caveat of longer training times.

>>>>

We conduct quantitative experiments to measure how different convolutional neural networks perceive low-level visual encodings, such as positions, angles, curvatures, and lengths. We formulate measurement tasks as logistic regression problems, where the networks must estimate single quantities or ratio relations between multiple quantities. The output between $0$ and $1$ indicates the percent of the value to be measured in the individual experiment.

% The output between $0$ and $1$ indicates the percent of the value to be measured in the individual experiment.

As baseline, we use a standard multilayer perceptron (MLP) without the generation of additional feature maps through convolutions (Fig.~\ref{fig:classifiers}).
For our experiments, we use a series of single factor between-subject designs with the factor \emph{network} (MLP, LeNet, VGG19, Xception, VGG19+ImageNet, Xception+ImageNet). We train each network in a supervised fashion, with each experiment including a training and a testing phase on data. without any overlap (Section~\ref{sec:data}). All parameters such as network hyperparameters, optimization methods, and stopping conditions are fixed (Section~\ref{sec:classifiers}). However, since the networks are of different complexity, the number of trainable parameters changes (Table~\ref{tab:parameters}). We define a series of hypotheses prior to each experiment.

@@

\subsection{Data}
\label{sec:data}
We create all visualizations as parametrized rasterized images without interpolation (except for the anti-alias experiment). The number of parameters differs per experiment as summarized in Table~\ref{tab:encoding_parameters} and section~\ref{sec:parametrizations}. 
\\~\\
\noindent\textbf{Noise.} We add subtle random noise ($0.05$) to each pixel to introduce additional variation. This prevents the classifiers of learning the actual pixels.
\\~\\
\noindent\textbf{Training/Validation/Test Splits.} We specify the size of each split set as follows: 60,000 training images, 20,000 validation images, and 20,000 test images. We then randomly add parameterized visualizations to the sets while guaranteeing that each set is disjunct from each other in terms of encoded variables. This eliminates leakage during training and evaluation. We also scale each set independently: images to the range of $-0.5$ to $0.5$ and labels to the range of $0.0$ to $1.0$. 
\\~\\
\noindent\textbf{Cross Validation.} For reproducibility, we perform repeated random sub-sampling validation, also known as Monte Carlo cross-validation, during our experiments. We run every experiment seperately multiple times (at least 4, up to 12) and randomly select (without replacement) the $60\%$ of our data as training data, $20\%$ as validation, and $20\%$ as test. Our large data sample size of 100,000 guarantees that every single observation of our parameterizations will be selected at least once (excluding noise patterns). Finally, we average the results over the runs.
\\~\\
\noindent\textbf{Intra-classifier Variability.} We also evaluate classifiers previously trained with one visualization on the same type of visualizations with different parameters by decreasing and increasing the variability of the generated images.
\\~\\
\noindent\textbf{Multi Classifiers.} Some experiments compare different types of visual encodings. We first train and evaluate each classifier on one type of encoding but we also train the classifiers on a random selection of multiple different types. These scenarios afflict the optimization process to find global minima and result in networks with more flexible knowledge with the caveat of longer training times.
